---
title: "lab_08"
author: "derek willis"
date: "8/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("174317d189a39d41b0ca921dd80d3965c5086e3f", overwrite=TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
ACS_variables <- load_variables(2019, "acs5", cache = TRUE)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
ppp_loans <-read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first.
**A1** 35.4

```{r}
md <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2019)

md <- md %>% arrange(desc(estimate)) 
```

**Q2** Which Maryland counties have a median age of at least 50?
**A2** Talbot and Worcester

```{r}
md %>%
  filter(estimate>=50)
  
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? Start with the code you wrote for Q4 and add a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** Prince George and Montgomery County.

```{r}

  
spanish_home <- get_acs(geography = "county",
              variables = c(medage = "B16004_004"),
              state = "MD",
              year = 2019)

total_population <- get_acs(geography = "county",
              variables = c(medage = "B07001_003"),
              state = "MD",
              year = 2019)

total_population %>%
  inner_join(spanish_home, by=c("GEOID", "NAME"))  %>%
  rename(poptotal=estimate.x, spantotal=estimate.y) %>%
  select(-variable.x, -variable.y) %>%
  mutate(spanish_percentage =((spantotal/poptotal)*100)) %>%
  arrange(desc(spanish_percentage))
  
  

  
  
```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4**
It's interesting that Montgomery County has a larger population total, but less spanish speakers. Out of the top five counties, Prince George and Montgomery have a vastly larger at-home spanish speaking population, but this is also most likely due to the larger populations both counties have. If I were to write about Garrett county, I would describe them as having a low diversity and low bi-literacy rate in their county, due to the less than 1 percent percentage of spanish-speakers at home. 

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** 60620

```{r}

black_population <-get_acs(geography="zcta", variables="B01001B_001", state=NULL)

total_population <- get_acs(geography="zcta", variables="B01003_001", state=NULL)

combined_population <-total_population %>%
  inner_join(black_population, by=c("GEOID", "NAME")) %>%
  rename(totalpop=estimate.x, blackpop=estimate.y)%>%
  select(-variable.x, -variable.y) %>%
  mutate(black_percent =((blackpop/totalpop)*100))

us_zipcodes <- ppp_loans %>%
  mutate(zip5 = str_sub(zip, 1, 5))

zip_loans <- us_zipcodes %>%
 group_by(zip5) %>%
 summarise(count=n()
           ) %>%
 arrange(desc(count))

final_zip <- zip_loans %>% left_join(combined_population, by=c('zip5'='GEOID'))


  

 
  


```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed application in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6**
The majority of the top ten zipcodes are located in Illinois. They seem to have in common that they are all located near Chicago in a more urban area. The only zipcode that sticks out is 30349, which is located near College Park, Georgia and is a suburban area.

**Q7** What other columns from the PPP data could be useful in reporting on what happened to this removed loans?
**A7** Count would be helpful because if an area has more PPP loans, it would make sense that there are more records of removed ppp loans. I would also consider the total population as well, to try and account for the number of loans received. 
