---
title: "state_profile"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PPP State Profile

Your assignment is to explore a dataset of PPP loan applications for a particular state and to answer a series of questions that will help you summarize that state's loan applications. You will need to write R code to answer those questions. You also will bring in other data, including Census information, and you will be making graphics, including maps, to illustrate your findings.

The deliverable will be this R Markdown notebook and a data folder that you will receive with a state's loan application data. Place any other data you are using to answer the questions in the same data folder.

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this assignment.

```{r}
library(tidyverse)
library(tidycensus)
library(janitor)
library(rvest)
library(sf)
library(tigris)
library(ggthemes)

```


NOTE FROM SHANNON: Thank you for all your help this semester! I really did struggle through this class, but I appreciate how understanding and helpful you were through the entire semester. While this profile really isn't perfect by any means, I tried my best to take what I learned this semester and apply it. Thank you for teaching me the basics of coding and although it will probably never come naturally to me, if you weren't as kind, patient and always willing to help - I genuinely think I would have failed this class. You are so appreciated more than you know! Thank you for not giving up on me!!


## Initial Exploration


**A1.** 
Hennepin County in Minnesota currently holds the highest amount of loans at 51,670 followed by Ramsey County at 16,934 and Dakota at 12,610. Hennepin and Ramsey both appear to be more urban counties, with large industries and buildings, and booming inner city. Dakota appears to be more of an rural area with focuses on water and soil conservation. The county also is named after the Dakota Sioux tribes who inhabited the area. The minimum amount of loans in Minnesota currently stands at a loan for $445.00 in Aitkins County. The maximum loan amount in Minnesota is also held in Aitkins county at 56084.00. Hennepin County has the largest number of loans at 51670. This isn't very surprising, as Hennepin county is the state's most populated city.


```{r}
#load in Minnesota ppp data

mn_data <- read_rds("~/Documents/GitHub/data_journalism_2021_fall/State Profile/Data/mn_ppp.rds")

#grouping MN data by project_county_name and arranging to see sum of total loans per county

mn_data %>%
  group_by(project_county_name)%>%
  summarise(total_loans=n()) %>%
  arrange(desc(total_loans))

#finding minimum loan amount in MN

mn_data %>%
  group_by(project_county_name) %>%
  summarise(min(amount))

#finding maximum loan amount in MN

mn_data %>%
  group_by(project_county_name) %>%
  arrange(desc(amount)) %>%
  summarise(max(amount)) 

#finding total count of loans per county
mn_data %>%
  group_by(project_county_name) %>%
  count(total_loans=n()) %>%
  arrange(desc(n)) 

#finding average loan amount per county
mn_data %>%
  group_by(project_county_name) %>%
  arrange(desc(amount)) %>%
  summarise(mean(amount))
 
  
  
  
  


```


## Geographic Analysis

**Q2.** Write R code that examines geographic patterns for PPP loans in your state, using Census population information to calculate a per-capita figure for the state and counties and zip codes. Then, make a county map using ggplot showing the per-capita data and a zip code map showing the difference from the statewide per-capita figure. Describe the most interesting or newsworthy findings based on your exploration.

**A2.** Okay so my code got a little funky, but what I found from the top five zip codes was that it seems the most PPP loans are going to nicer/more urban areas. In the top 5 zip codes, the cities include Austin, Saint Paul, Burnsville, Hennepin County and Minneapolis - all of which are fairly nicer areas in MN and more populated. I think the most interesting data from this set is that those who are getting PPP loans in these areas probably have jobs that skew towards more of the typical business/office industry and are not necessarily receiving loans for rural farming industries. 
```{r}
#load in census_api_key
census_api_key("174317d189a39d41b0ca921dd80d3965c5086e3f", overwrite=TRUE)

#read in mn data
mn_data

#filter mn data to zip5
mn_data <- mn_data %>%
  mutate(zip5 = str_sub(zip, 1, 5))

mn_zip_data <- mn_data %>%
  group_by(zip5)%>%
  count(total_loan_amount=n())%>%
  arrange(desc(n))

#reading in ACS variables so that I can search for data by zip code
ACS_variables <- load_variables(2019, "acs5", cache = TRUE)

#pulling MN zipcode data from the us ppp dataset
mn_state_zip <-get_acs(geography = "zcta",
              variables = "B01001_001",
              state="MN",
              year = 2019)

#joining MN zipcode dataset with total mn_data zips
combined_mn_zips <- mn_data %>%
  inner_join(mn_state_zip, by=c('zip5'='GEOID'))
  

#pulling MN county data from total counties
counties <- counties()

mn_counties <- counties %>%
  filter(STATEFP == "27")


mn_counties %>%
  ggplot() + 
  geom_sf() + 
  theme_minimal()

ggplot() + 
  geom_sf(data=mn_state_zip) + 
  theme_minimal()

#creating geographic map on ggplot
ggplot() + 
  geom_sf(data=mn_state_zip, aes(fill=mn_state_zip)) + 
  theme_minimal() +
  scale_fill_viridis_b(option="magma",trans = "log")



```

## Lender Analysis

**Q3.** Write R code to examine which lenders had the most approved applications in your state (and include summary statistics such as total amount and average loan amount). Generate dataframes that show the number and total amount of all loans and undisbursed loans per lender. For those lenders who had any undisbursed loans, make a graphic showing the relationship between the total amount of loans and the total undisbursed amount. Describe the most noticeable outlier lenders on that graphic below.

**A3.** The Village Bank and Trust Association stands out as the largest outlier on the chart. The association only had 4 lenders, but had a large undisbursed amount of 4139700. After a little research, the organization's front page boasts that the bank provides small businesses with PPP loans, as well as help with PPP loan forgiveness. It's interesting that only 4 small businesses trusted the organization with their money, and that the bank never even ended up disbursing the money out. I wonder if the businesses ending up failing prior to the distrubtion of the money, or if the bank ended up not being super trustworthy.

```{r}
#creating new dataframe to hold undisbursed lender data and number of loans per lender
mn_lender_undisbursed_data <- mn_data %>%
  group_by(lender)%>%
  summarise(undisbursed_amount = sum(undisbursed_amount)) %>%
  arrange(desc(undisbursed_amount)) %>%
  head(10)


#finding total loans count per lender
lender_count <- mn_data %>%
  group_by(lender) %>%
  count(total_loans=n()) %>%
  arrange(desc(n))

#dataframe to hold disbursed lender data and # of loans per lender 
mn_disbursed_data <- mn_data %>%
  group_by(lender) %>%
  summarise(current_approval_amount= sum(current_approval_amount),
                  total_loans=n()) %>%
  arrange(desc(current_approval_amount))

#plotting undisbursed loans for top ten lenders 

mn_lender_undisbursed_data %>%
  ggplot() +
  geom_bar(aes(x=reorder(lender,undisbursed_amount), weight=total_loans)) +
  coord_flip() +
  theme_economist()+
  labs(
    title="Mn Undisbursed Lender Loans",
    x = "Lender Undisbursed Amount",
    y = "Total Loans",
    caption = "source: SBA PPP Loan Database"
    
  )
    


```


## Industry Analysis

**Q4.** Write R code that examines industry patterns for PPP loans in your state, using the NAICS codes from the PPP data as a starting point. Generate statewide and county industry totals, then join that with 2018-19 data from the [Census County Business Patterns survey](https://www2.census.gov/programs-surveys/cbp/datasets/2019/cbp19co.zip) using 6-digit NAICS codes. The documentation explaining that data can be found here: https://www2.census.gov/programs-surveys/cbp/technical-documentation/records-layouts/2018_record_layouts/county-layout-2018.txt. To do this, you will need to add FIPS codes to your PPP dataset.

Does the distribution of PPP applications by the top 10 industries (by number of applications) roughly match the number of businesses reported in the Census data? Does it roughly match if you remove self-employed individuals and sole proprietorships from the PPP totals? Write up a summary of what you've found and whether there are potentially newsworthy patterns in the data.

Create a county-level map showing the differences between the PPP applications and the Census business data for one of the top 10 industry codes. You can do this either using ggplot or Datawrapper (if the latter, publish that map and include the URL below).

**A4.** 
I'm sorry I just got so stuck on this part.

```{r}
#reading in new dataframe for MN businesses and filtering by top ten businesses

mn_businesses <- mn_data %>%
  group_by(naics_code)%>%
  summarise(naics_code=n()) %>%
  arrange(desc(naics_code)) %>%
  head(10)

#using naics code 6920, finding industry totals per county

mn_industry<- mn_data%>%
  filter(naics_code==6920) %>%
  group_by(project_county_name)





```

## Summary

**Q5.** What is the most promising story idea that you have found during this exploration? How would you approach reporting it out and what other information/data would you need to do that? What would be the minimum story you could get from this, and what would be the maximum story?

**A5.** I think the most promising story idea I have found during this exploration lies in the industry data. There is a strange amount of undisbursed laon amounts, among seemingly low numbers of loans in certain counties. To further my reporting, I would specfially like to look into the outliers in the data - starting with Village Bank as well as Capital Plus Financial, to look at the differences in the loan approval process, as well as the specifics behind the financial institution. I would also look into the area these businesses are located in, focusing in on the demographic analysis in these areas. Do these institutions tend to approve loans and target certain people groups? Are they located in rural or urban areas and how often do they approve loans? Then coming in contact with the people who were approved for these loans would be the next step to hear their stories about the loan process, as well as conducting interviews with individuals who work at the organizations. I think the minimum story I could get from this would just fall under something such as - not a ton of businesses applied in the area for SBA loans from a specfic bank and they never ended up using the loans (or ended up being faulty organizations.) The maximum story could fall under something such as loan approval bias, or fradulent/discrimintatory behavior from the lenders. 
